---
title: "Lauren Latham - Lab 02 - Plastic waste"
author: "laurenlatham97"
date: "2/4/21"
output: github_document
---

## Load packages and data

```{r load-packages, message=FALSE}
library(tidyverse) 
```

```{r load-data, message=FALSE}
plastic_waste <- read_csv("data/plastic-waste.csv")
```

Below is a histogram of the distribution of plastic waste per capita faceted by continent. AND WHAT DOES THIS SAY ABOUT HOW COUNTRIES COMPARE? 

```{r}
ggplot(plastic_waste, aes(x=plastic_waste_per_cap))+geom_histogram(bins=15)+facet_wrap(~continent)
```

Below is a density plot demonstrating plastic waste per capita with each continent color-coded. I lowered the alpha to 0.3 so that data from each continent can be seen more easily. The color and fill go in the mapping aesthetics because they are mapped to specific variables, whereas the alpha information goes in the geom characteristics because it is information that is used in the same way for each variable (i.e., each continent). 
```{r}
ggplot(plastic_waste, aes(x=plastic_waste_per_cap, color=continent, fill=continent))+geom_density(alpha=.3)
```

Below are side by side box plots displaying plastic waste per capita for each continent: 

```{r}
ggplot(data = plastic_waste, 
       mapping = aes(x = continent, 
                     y = plastic_waste_per_cap)) +
  geom_boxplot()
```

And violin plots...


```{r}
ggplot(data = plastic_waste, 
       mapping = aes(x = continent, 
                     y = plastic_waste_per_cap)) +
  geom_violin()
```

The violin plots are useful because the display the full distribution of the data, unlike the box plots. OK BUT WHAT DO BOX PLOTS DISPLAY THAT VIOLIN PLOTS DO NOT? - REVISIT.


Relationship between plastic waste per capita and mismanaged plastic waste per capita (scatterplot):

```{r}
ggplot(data = plastic_waste, 
       mapping = aes(x = plastic_waste_per_cap, 
                     y = mismanaged_plastic_waste_per_cap)) +
  geom_point()
```

From the above scatterplot we can see that there is a positive relationship between plastic waste per capita and mismanaged plastic wast per capita (i.e., as plastic waste per capita goes up, mismanaged plastic waster per capita also goes up). However, it looks like there may be two slightly different patterns, with some countries having a very strong positive relationship between these variables, and other countries have a somewhat weaker positive relationship. 

Relationship between plastic waste per capita and mismanaged plastic waste per capita (scatterplot-color country): 

```{r}
ggplot(data = plastic_waste, 
       mapping = aes(x = plastic_waste_per_cap, 
                     y = mismanaged_plastic_waste_per_cap, color = continent)) +
  geom_point()
```

The above scatterplot demonstrates the same relationship described earlier between plastic waste per capita and mismanaged plastic waste per capita, but the color coding makes it easier to understand the two potentially different patterns. It looks like the relationship between these variables is weaker for North American and European countries, and much stronger for other countries. 

Relationship between total population and plastic waste per capita (scatterplot):
```{r}
ggplot(data = plastic_waste, 
       mapping = aes(x = total_pop, 
                     y = plastic_waste_per_cap, color = continent)) +
  geom_point()
```
INSERT EXPLANATION

Relationship between coastal population and plastic waste per capita (scatterplot): 
```{r}
ggplot(data = plastic_waste, 
       mapping = aes(x = coastal_pop, 
                     y = plastic_waste_per_cap, color = continent)) +
  geom_point()
```
INSERT EXPLANATION
