---
title: "Lab 02 - Plastic waste"
author: "laurenlatham97"
date: "2/4/21"
output: github_document
---

## Load packages and data

```{r load-packages, message=FALSE}
library(tidyverse) 
```

```{r load-data, message=FALSE}
plastic_waste <- read_csv("data/plastic-waste.csv")
```

Below is a histogram of the distribution of plastic waste per capita faceted by continent. WHAT DOES THIS SAY ABOUT HOW CONTINENTS COMPARE TO EACHOTHER??

```{r}
ggplot(plastic_waste, aes(x=plastic_waste_per_cap))+geom_histogram(bins=15)+facet_wrap(~continent)
```

Below is a density plot demonstrating plastic waste per capita with each continent color-coded. I lowered the alpha to 0.3 data from each continent can be seen more easily. The color and fill go in the mapping aesthetics because they are mapped to specific variables, whereas the alpha information goes in the geom characteristics because it is information that is used in the same way for each variable (i.e., each continent). 
```{r}
ggplot(plastic_waste, aes(x=plastic_waste_per_cap, color=continent, fill=continent))+geom_density(alpha=.3)
```

Below are side by side box plots displaying plastic waste per capita for each continent: 

```{r}
ggplot(data = plastic_waste, 
       mapping = aes(x = continent, 
                     y = plastic_waste_per_cap)) +
  geom_boxplot()
```

And violin plots...

```{r}
ggplot(data = plastic_waste, 
       mapping = aes(x = continent, 
                     y = plastic_waste_per_cap)) +
  geom_violin()
```

The violin plots are useful because the display the full distribution of the data, unlike the box plots. OK BUT WHAT DO BOX PLOTS DISPLAY THAT VIOLIN PLOTS DO NOT? - REVISIT.

scatterplot:

```{r}
ggplot(data = plastic_waste, 
       mapping = aes(x = plastic_waste_per_cap, 
                     y = mismanaged_plastic_waste_per_cap)) +
  geom_point()
```
scatterplot-color: 

```{r}
ggplot(data = plastic_waste, 
       mapping = aes(x = plastic_waste_per_cap, 
                     y = mismanaged_plastic_waste_per_cap, color = continent)) +
  geom_point()
```

```{r}
ggplot(data = plastic_waste, 
       mapping = aes(x = plastic_waste_per_cap, 
                     y = total_pop, color = continent)) +
  geom_point()
```

```{r}
ggplot(data = plastic_waste, 
       mapping = aes(x = plastic_waste_per_cap, 
                     y = coastal_pop, color = continent)) +
  geom_point()
```

Still need to work on fixing up these graphs and making interpretations as instructed in the lab, but the base code is all here. 