---
title: "Lauren Latham - Lab 02 - Plastic waste"
author: "laurenlatham97"
date: "2/4/21"
output: github_document
---

## Load packages and data

```{r load-packages, message=FALSE}
library(tidyverse) 
```

```{r load-data, message=FALSE}
plastic_waste <- read_csv("data/plastic-waste.csv")
```
### Exercise 1 - distribution of plastic waste faceted by continent

Below is a histogram of the distribution of plastic waste per capita faceted by continent. The histogram reveals that North America contains the countries that have the most plastic waste per capita. Additionally, the distribution seems to be positively skewed for most continents. 

```{r plot-plastic_waste, fig.height=8, fig.width=8}
ggplot(plastic_waste, aes(x=plastic_waste_per_cap,))+geom_histogram(bins=30)+facet_wrap(~continent)
```

### Exercise 2 - density plot of pastic waste per capita

Below is a density plot demonstrating plastic waste per capita with each continent color-coded. I lowered the alpha to 0.3 so that data from each continent can be seen more easily. The color and fill go in the mapping aesthetics because they are mapped to specific variables, whereas the alpha information goes in the geom characteristics because it is information that is used in the same way for each variable (i.e., each continent). 

```{r plot-pastic_waste 2, fig.height=8, fig.width=8}
ggplot(plastic_waste, aes(x=plastic_waste_per_cap, color=continent, fill=continent))+geom_density(alpha=.3)
```

### Exercise 3 - box plots and violin plots
Below are side by side box plots displaying plastic waste per capita for each continent: 

```{r plot-pastic_waste 3, fig.height=8, fig.width=8}
ggplot(data = plastic_waste, 
       mapping = aes(x = continent, 
                     y = plastic_waste_per_cap)) +
  geom_boxplot()
```

And violin plots...


```{r plot-pastic_waste 4, fig.height=8, fig.width=8}
ggplot(data = plastic_waste, 
       mapping = aes(x = continent, 
                     y = plastic_waste_per_cap)) +
  geom_violin()
```

The violin plots are useful because the display the full distribution of the data, unlike the box plots. For example, a bimodal distribution is very easily obscured by box plots. Box plots however make it easier to discern certain summary statistics (median, upper and lower quartile, outliers). 

### Exercise 4 - scatterplots
Relationship between plastic waste per capita and mismanaged plastic waste per capita (scatterplot):

```{r plot-pastic_waste 5, fig.height=8, fig.width=8}
ggplot(data = plastic_waste, 
       mapping = aes(x = plastic_waste_per_cap, 
                     y = mismanaged_plastic_waste_per_cap)) +
  geom_point()
```

From the above scatterplot we can see that there is a positive relationship between plastic waste per capita and mismanaged plastic wast per capita (i.e., as plastic waste per capita goes up, mismanaged plastic waster per capita also goes up). However, it looks like there may be two slightly different patterns, with some countries having a very strong positive relationship between these variables, and other countries have a somewhat weaker positive relationship. 

Relationship between plastic waste per capita and mismanaged plastic waste per capita (scatterplot-color country): 

```{r plot-pastic_waste 6, fig.height=8, fig.width=8}
ggplot(data = plastic_waste, 
       mapping = aes(x = plastic_waste_per_cap, 
                     y = mismanaged_plastic_waste_per_cap, color = continent)) +
  geom_point()
```

The above scatterplot demonstrates the same relationship described earlier between plastic waste per capita and mismanaged plastic waste per capita, but the color coding makes it easier to understand the two potentially different patterns. It looks like the relationship between these variables is weaker for North American and European countries, and much stronger for other countries. 

Relationship between total population and plastic waste per capita (scatterplot):
```{r plot-pastic_waste, fig.height=8, fig.width=8}
ggplot(data = plastic_waste, 
       mapping = aes(x = total_pop, 
                     y = plastic_waste_per_cap, color = continent)) +
  geom_point()

```

Relationship between coastal population and plastic waste per capita (scatterplot): 
```{r plot-pastic_waste 7, fig.height=8, fig.width=8}
ggplot(data = plastic_waste, 
       mapping = aes(x = coastal_pop, 
                     y = plastic_waste_per_cap, color = continent)) +
  geom_point()
```

Neither total population nor coastal population appear to be strongly linearly associated with plastic waste per capita. It's somewhat difficult to tell however, because many of the countries cluster around the same total population and coastal population. Interestingly, a lot of the Asian countries with very large coastal populations seem to be relatively low in plastic waster per capita, this is less true however for total population. The Oceanic and NOrth American countries however with lower coastal population and total population seem to fall pretty high in plastic waste per capita. 
